---
title: "Tutorial VISA"
author: "Thomas Goossens"
date: "3/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r pkg-config, eval = FALSE}
install.packages("packrat")
packrat::init()
install.packages("devtools")
devtools::install_github("pokyah/mlr", ref = "gstat")
install.packages("sf")
install.packages("sp")
install.packages("mapview")
install.packages("leaflet.extras")
install.packages("dplyr")
install.packages("gstat")
install.packages("automap")
install.packages("RColorBrewer")
install.packages("bestNormalize")
```

```{r data-load}
dataset = read.csv(file = "./raw-data/tracteur.csv", header = TRUE,sep = ";" )
shapefile = sf::st_read(dsn = "./raw-data/shapefile/reves_WGS84.shp")
```

```{r map-shapefile}
mapview::mapview(shapefile)
```

```{r map-records}
# transforming to spatial sf object
dataset.sf = sf::st_as_sf(dataset, coords = c("long_deg", "lat_deg"))

# setting the CRS (coordinate reference system)
dataset.sf = sf::st_set_crs(dataset.sf, 4326)

# map the patinage & traction on static map
patinage.map = plot(dataset.sf["patinage"])
traction.map = plot(dataset.sf["traction"])
```

A point seems to be out of the parcel. Lets remove it by intersecting with the parcel. To do this both objects must be in the same CRS (which is the case - see above).

```{r records-in-parcel}
# instersecting
dataset.sf = sf::st_intersection(dataset.sf, shapefile)

# checking the new maps
patinage.map = plot(dataset.sf["patinage"])
traction.map = plot(dataset.sf["traction"])
```

Let's construct a grid of pixels on which we will interpolate

```{r interpolation-grid}
# Building a grid of cells of 10 m². To do so, we must work with projected data (to define the cellsize)
interpolation.grid.sf = sf::st_make_grid(x = sf::st_transform(shapefile, crs = 3812),  cellsize = 10, what = "centers")

# mapit on an interactive map 
mapview::mapview(interpolation.grid.sf)
```

```{r machine-learning-tasks-preparation}
library(dplyr)

# converting the spatial object to normal dataframe
dataset = dataset.sf

# making it projected instead of coordinates (for inteprolation)
dataset = sf::st_transform(dataset, 3812)
coordinates = data.frame(sf::st_coordinates(dataset))
sf::st_geometry(dataset) = NULL

# putting coordinates in the dataframe
dataset = dataset %>%
  dplyr::bind_cols(coordinates)

# renaming X and Y to x and y for gstat learner to work
dataset = dataset %>%
  dplyr::rename(x = X) %>%
  dplyr::rename(y = Y)

# creating the 2 tasks
traction.regr.task = mlr::makeRegrTask(
  data = dataset, target = "traction") 
patinage.regr.task = mlr::makeRegrTask(
  data = dataset, target = "patinage")
```

```{r machine-learning-algos}
learners = list(
  # linear regression
  LR = mlr::makeFilterWrapper(
      learner = mlr::makeLearner(
        cl = "regr.lm",
        id = "multiReg.x_y",
        predict.type = 'se'),
      fw.method = "linear.correlation",
      fw.mandatory.feat = c("y", "x"),
      fw.abs = 2),
  # inverse distance weigthed
  IDW = mlr::makeLearner(
    cl = "regr.gstat",
    id = "idw"),
  # trend surface order 1, 2 and 3
  TS1 = mlr::makeLearner(cl = "regr.gstat", id = "mlr-ts1", degree = 1),
  TS2 = mlr::makeLearner(cl = "regr.gstat", id = "mlr-ts2", degree = 2),
  TS3 = mlr::makeLearner(cl = "regr.gstat", id = "mlr-ts3", degree = 3)
)

```

```{r algos-benchmarking}
# setting the seed for reprocucible results
set.seed(86)

# defining how we split between training and validation set
rdesc = mlr::makeResampleDesc("CV", iters = 5)

# benchmark for traction
traction.bmr = mlr::benchmark(
  learners = learners, resamplings = rdesc, tasks = traction.regr.task, measures = list(mlr::rmse, mlr::mse, mlr::mae))

# plot summary of bmr
mlr::plotBMRSummary(traction.bmr, pretty.names = FALSE)

# resample(learner = learner.rf, task = spatial.task,
#   resampling = resampling, measures = list(auc))

```
TS3 is the best predictor (if we ignore IDW). Let's use this to spatialize on our grid

```{r train-learner-TS3-IDW}
model.TS3 = mlr::train(learners$TS3, traction.regr.task)
model.IDW = mlr::train(learners$IDW, traction.regr.task)
```

```{r make-grid-pred}
# converting the interpolation grid to normal dataframe
interpolation.grid = interpolation.grid.sf
# making it projected instead of coordinates
interpolation.grid = sf::st_transform(interpolation.grid, 3812)
coordinates = data.frame(sf::st_coordinates(interpolation.grid))

# putting coordinates in the dataframe
interpolation.grid = coordinates

# renaming X and Y to x and y for gstat learner to work
interpolation.grid  = interpolation.grid  %>%
  dplyr::rename(x = X) %>%
  dplyr::rename(y = Y)

# using the model to compute the predictions
predictions.TS3 = predict(model.TS3, newdata = interpolation.grid)$data
predictions.IDW = predict(model.IDW, newdata = interpolation.grid)$data

# Building a grid of pixels of 10 m² that spatially matches our grid of centers. To do so, we must work with projected data (to define the cellsize)
interpolation.grid.pixels.sf = sf::st_sf(sf::st_make_grid(x = sf::st_transform(shapefile, crs = 3812),  cellsize = 10, what = "polygons"))

# adding the predictions to the grid squares object
spatialized.TS3.sf = interpolation.grid.pixels.sf %>%
        dplyr::bind_cols(predictions.TS3)

spatialized.IDW.sf = interpolation.grid.pixels.sf %>%
        dplyr::bind_cols(predictions.IDW)

# making it a spatial object again for mapping
#spatialized.TS3.sf = sf::st_as_sf(spatialized.TS3, coords = c("y", "x"))
# setting the CRS
#spatialized.TS3.sf = sf::st_set_crs(spatialized.TS3.sf, 3812)

# limit it to our parcel shapefile and not full extent
spatialized.TS3.sf = sf::st_intersection(spatialized.TS3.sf, sf::st_transform(shapefile, crs = 3812))

spatialized.IDW.sf = sf::st_intersection(spatialized.IDW.sf, sf::st_transform(shapefile, crs = 3812))

# mapping 
mapview::mapview(spatialized.TS3.sf["response"])
plot(spatialized.TS3.sf["response"])
plot(spatialized.IDW.sf["response"])
```

Now let's try with some Kriging

```{r kriging-gstat-way}
# doing kriging the gstat way for explorative purposes
# making spatial sp objects
dataset.sp = dataset
interpolation.grid.sp = interpolation.grid
sp::coordinates(dataset.sp) = ~x+y
sp::coordinates(interpolation.grid.sp) = ~x+y
sp::gridded(interpolation.grid.sp) = TRUE

# to solve the warnings while using gstat::krige => https://gis.stackexchange.com/questions/222192/r-gstat-krige-covariance-matrix-singular-at-location-5-88-47-4-0-skipping

#::FIXME:: better solution to remove duplicate locations
dataset.sp = dataset.sp[-sp::zerodist(dataset.sp)[,1],] 

# making and plotting the variogram
traction.vgm = gstat::variogram(traction~1, dataset.sp)
plot(traction.vgm)

# manually fitting a model to the vgm with constant mean
# https://gisgeography.com/semi-variogram-nugget-range-sill/
traction.fit = gstat::fit.variogram(traction.vgm, model = gstat::vgm(
  psill = max(traction.vgm$gamma), model = "Sph", range =  150, nugget = 5000000))
plot(traction.vgm, traction.fit)

# interpolating
# to solve the memory error  => https://github.com/edzer/gstat/issues/6
traction.kr = gstat::krige(traction~1, dataset.sp, interpolation.grid.sp, model = traction.fit, maxdist = 30)

# plotting the interpolation
traction.kr.plot = sp::spplot(traction.kr["var1.pred"], do.log = F, colorkey = TRUE)
traction.kr.plot
```

```{r kriging-mlr-way}
# reusing the cleaned dataset
dataset = as.data.frame(dataset.sp)

# making the task
traction.regr.task = mlr::makeRegrTask(
  data = dataset, target = "traction") 

# defining the kriging learner based on the parameter found via the variogram
KR = mlr::makeFilterWrapper(
    learner = mlr::makeLearner(
      cl = "regr.gstat",
      id = "ok",
      par.vals = list(
        range = 150,
        psill = max(traction.vgm$gamma),
        model.manual = "Sph",
        nugget = 5000000,
        maxdist = 50,
        debug.level = 0),
      predict.type = "se"),
    fw.method = "linear.correlation",
    fw.mandatory.feat = c("y", "x"),
    fw.abs = 2)

# adding this learner to our learners list
learners = c(learners, list(KR = KR))

# redoing the bmr with the new KR learner
# traction.bmr = mlr::benchmark(
#   learners = learners, resamplings = rdesc, tasks = traction.regr.task, measures = list(mlr::rmse))

# plot summary of bmr
mlr::plotBMRSummary(traction.bmr, pretty.names = FALSE)

# training the learner to build the model
model.KR = mlr::train(learner = learners$KR, task = traction.regr.task)

# using the model to compute the predictions
predictions.KR = predict(model.KR, newdata = interpolation.grid)$data

# adding the predictions to the grid squares object
spatialized.KR.sf = interpolation.grid.pixels.sf %>%
        dplyr::bind_cols(predictions.KR)

# making it a spatial object again for mapping
# spatialized.KR.sf = sf::st_as_sf(spatialized.KR, coords = c("y", "x"))
# # setting the CRS
# spatialized.KR.sf = sf::st_set_crs(spatialized.KR.sf, 3812)

# limit it to our parcel shapefile and not full extent
spatialized.KR.sf = sf::st_intersection(spatialized.KR.sf, sf::st_transform(shapefile, crs = 3812))

# mapping 
mapview::mapview(spatialized.KR.sf["response"])
plot(spatialized.KR.sf["response"])


# redoing the benchmark for traction using all the learners
# traction.bmr = mlr::benchmark(
#   learners = learners, resamplings = rdesc, tasks = traction.regr.task, measures = list(mlr::rmse))
```


```{r train-learner-KR}
model.KR = mlr::train(learners$KR, traction.regr.task)
```

```{r make-grid-pred-2}
# using the model to compute the predictions
predictions.KR = predict(model.KR, newdata = interpolation.grid)$data

# adding the predictions to the grid squares object
spatialized.KR = interpolation.grid.pixels.sf %>%
        dplyr::bind_cols(predictions.KR)

# making it a spatial object again for mapping
spatialized.KR.sf = sf::st_as_sf(spatialized.KR, coords = c("y", "x"))
# setting the CRS
spatialized.KR.sf = sf::st_set_crs(spatialized.KR.sf, 3812)

# limit it to our parcel shapefile and not full extent
spatialized.KR.sf = sf::st_intersection(spatialized.KR.sf, sf::st_transform(shapefile, crs = 3812))

# mapping 
mapview::mapview(spatialized.KR.sf["response"])
plot(spatialized.KR.sf["response"])
```

Making a better map with leaflet

```{r leaflet}
library(leaflet)

# to make the map responsive
responsiveness = "\'<meta name=\"viewport\" content=\"width=device-width, initial-scale=1.0\">\'"
  
# defining the color palette for traction
tractionPal = leaflet::colorNumeric(
  palette = "RdYlBu", #"RdBl",
  reverse = TRUE,
  domain = spatialized.KR.sf$response, #spatialized$response,
  na.color = "black"
)

# projecting in WGS84
spatialized.KR.sf = sf::st_transform(spatialized.KR.sf, 4326)

traction.map = leaflet::leaflet(spatialized.KR.sf) %>%
    # basemaps
    leaflet::addProviderTiles(group = "Satellite",
      providers$Esri.WorldImagery,
      options = providerTileOptions(opacity = 1)
    ) %>%
    leaflet::addProviderTiles(group = "Stamen",
      providers$Stamen.Toner,
      options = providerTileOptions(opacity = 0.25)
    ) %>%
    # centering the map
    leaflet::fitBounds(sf::st_bbox(spatialized.KR.sf)[[1]],
      sf::st_bbox(spatialized.KR.sf)[[2]],
      sf::st_bbox(spatialized.KR.sf)[[3]],
      sf::st_bbox(spatialized.KR.sf)[[4]]
    ) %>%
    # adding layer control button
    leaflet::addLayersControl(
      baseGroups = c("Satellite","Stamen"),
     # overlayGroups = c("prediction"),
      overlayGroups = c("response"),
        options = layersControlOptions(collapsed = TRUE)
      ) %>%
    # fullscreen button
    leaflet.extras::addFullscreenControl() %>%
    # location button
    leaflet::addEasyButton(easyButton(
      icon = "fa-crosshairs", title = "Locate Me",
      onClick = JS("function(btn, map){ map.locate({setView: true}); }"))) %>%
    htmlwidgets::onRender(paste0("
      function(el, x) {
      $('head').append(",responsiveness,");
      }")
    ) %>%
    # admin boundaries
    # predictions
    leaflet::addPolygons(
      group = "response",
      color = "#444444", stroke = FALSE, weight = 1, smoothFactor = 0.8,
      opacity = 1.0, fillOpacity = 0.6,
      fillColor = ~tractionPal(response),
      label = ~ paste(
        "prediction:", format(round(spatialized.KR.sf$response, 2), nsmall = 2)),
      # popup = ~as.character(response),
      highlightOptions = highlightOptions(color = "white", weight = 2,
        bringToFront = TRUE)
    ) %>%
    #legend
    leaflet::addLegend(
      position = "bottomright", pal = tractionPal, values = ~response,
      title = "traction",
      group = "response",
      opacity = 1
    )


```

Normalization

```{r reading-data-raw}
dataset = read.csv(file = "./raw-data/tracteur.csv", header = TRUE,sep = ";" )
shapefile = sf::st_read(dsn = "./raw-data/shapefile/reves_WGS84.shp")

# transforming to spatial sf object
dataset.sf = sf::st_as_sf(dataset, coords = c("long_deg", "lat_deg"))
# transforming to spatial sf object
dataset.sf = sf::st_as_sf(dataset, coords = c("long_deg", "lat_deg"))
# setting the CRS (coordinate reference system)
dataset.sf = sf::st_set_crs(dataset.sf, 4326)
# instersecting
dataset.sf = sf::st_intersection(dataset.sf, shapefile)

# solution to remove duplicate locations
dataset.sp = as(dataset.sf, "Spatial")[-sp::zerodist(as(dataset.sf, "Spatial"))[,1],] 
dataset.sf = sf::st_as_sf(dataset.sp)

# making it projected instead of coordinates (for interpolation) and tranforming to dataframe
dataset = sf::st_transform(dataset.sf, 3812)
coordinates = data.frame(sf::st_coordinates(dataset))
sf::st_geometry(dataset) = NULL

# putting coordinates in the dataframe
dataset = dataset %>%
  dplyr::bind_cols(coordinates)

# renaming X and Y to x and y for gstat learner to work
dataset = dataset %>%
  dplyr::rename(x = X) %>%
  dplyr::rename(y = Y)

# plotting the map
plot(sf::st_as_sf(dataset, coords = c("y","x"))["traction"])
```


```{r histogram}
# traction.d = density(dataset$traction)
# plot(traction.d)
# traction.n = scale(dataset$traction)
# traction.n.d = density(traction.n)
# plot(traction.n.d)

# normalization with boxcox
# https://github.com/petersonR/bestNormalize
dataset = dataset %>%
  dplyr::mutate(traction_m = traction + abs(min(traction)) + 1)

traction.bcn = bestNormalize::boxcox(dataset$traction_m)

traction.bcn.d = density(traction.bcn$x.t)
plot(traction.bcn.d)

dataset = dataset %>%
  dplyr::bind_cols(traction.bcn = data.frame(traction.bcn$x.t)) %>%
  dplyr::rename(traction.norm = traction.bcn.x.t)

# plot the map
plot(sf::st_as_sf(dataset, coords = c("y","x"))["traction.norm"])
```

```{r kriging after normalization and residuals}
# defining the poly2 plan model
traction.poly2 = spatial::surf.ls(np = 2, x = dataset$x, y = dataset$y, z = dataset$traction.norm)

# poly2 predictions at observations
traction.poly2.preds = predict(object = traction.poly2, x = dataset$x, y = dataset$y)

# data - trend for residuals
traction.poly2.res = dataset$traction.norm - traction.poly2.preds

# adding to dataset
dataset = dataset %>%
  dplyr::bind_cols(traction.norm.res = traction.poly2.res )

# plot the map
plot(sf::st_as_sf(dataset, coords = c("y","x"))["traction.norm.res"])

# dataset as spatial object class sp
dataset.sf = dataset
dataset.sf = sf::st_as_sf(dataset.sf, coords = c("y", "x"))
dataset.sf = sf::st_set_crs(dataset.sf, 3812)
coordinates = data.frame(sf::st_coordinates(dataset.sf))

# putting coordinates in the dataframe ::FIXEME:: maybe useless
dataset.sf = dataset.sf %>%
  dplyr::bind_cols(coordinates)

# renaming X and Y to x and y for gstat learner to work
dataset.sf = dataset.sf %>%
  dplyr::rename(x = X) %>%
  dplyr::rename(y = Y)

plot(dataset.sf["traction.norm.res"] )
dataset.sp = as(dataset.sf, "Spatial")
sp::spplot(dataset.sp["traction.norm.res"])

#::ALLGOOD::#

# making and plotting the variogram
traction.norm.res.vgm = gstat::variogram(traction.norm.res~1, dataset.sp)
plot(traction.norm.res.vgm)

# manually fitting a model to the vgm with constant mean
# https://gisgeography.com/semi-variogram-nugget-range-sill/
traction.norm.res.fit = gstat::fit.variogram(
  object = traction.norm.res.vgm,
  model = gstat::vgm(
    psill = 0.9, model = "Sph", range =  150, nugget = 0.7))
plot(traction.norm.res.vgm, traction.norm.res.fit)

# remaking the grid from scratch
interpolation.grid.sf = sf::st_make_grid(x = sf::st_transform(shapefile, crs = 3812),  cellsize = 10, what = "centers")

# converting the interpolation grid to normal dataframe
interpolation.grid = interpolation.grid.sf
# making it projected instead of coordinates
interpolation.grid = sf::st_transform(interpolation.grid, 3812)
coordinates = data.frame(sf::st_coordinates(interpolation.grid))

# putting coordinates in the dataframe
interpolation.grid = coordinates

# renaming X and Y to x and y for gstat learner to work
interpolation.grid  = interpolation.grid  %>%
  dplyr::rename(x = X) %>%
  dplyr::rename(y = Y) %>%
  dplyr::bind_cols(coordinates)

interpolation.grid.sf = sf::st_as_sf(interpolation.grid, coords = c("Y", "X"))
interpolation.grid.sf = sf::st_set_crs(interpolation.grid.sf, 3812)

interpolation.grid.sp = as(interpolation.grid.sf, "Spatial")
sp::gridded(interpolation.grid.sp) = TRUE


# interpolating the residuals
# to solve the memory error  => https://github.com/edzer/gstat/issues/6
# https://gis.stackexchange.com/questions/245295/error-in-kriging-function
traction.norm.res.kr = gstat::krige(
  traction.norm.res~1,
  dataset.sp,
  interpolation.grid.sp,
  model = traction.norm.res.fit,
  maxdist = 30)

traction.norm.res.kr.sf = sf::st_as_sf(traction.norm.res.kr)

# plotting the interpolation of the residuals
traction.norm.res.kr.plot = sp::spplot(traction.norm.res.kr["var1.pred"], do.log = F, colorkey = TRUE)
traction.norm.res.kr.plot

# poly2 predictions at prediction grid points locations
traction.poly2.grid.preds = predict(
  object = traction.poly2,
  x = data.frame(interpolation.grid.sp)$x,
  y = data.frame(interpolation.grid.sp)$y)

# Addition of prediction of kriging of residuals and polynomial2 trend predictions at grid points locations
traction.norm.kr = data.frame(traction.norm.res.kr["var1.pred"]) + traction.poly2.grid.preds

# denormalizing the data by performing reverse transformation
traction.kr =  predict(traction.bcn, newdata = traction.norm.kr$var1.pred, inverse = TRUE)
traction.kr = data.frame(traction.kr = traction.kr)

# to get back to the original data
traction.kr = traction.kr %>%
    dplyr::mutate(traction.kr = traction.kr - (abs(min(dataset$traction))) - 1)

# joining to the spatial object
spatialized.kr = interpolation.grid %>%
  dplyr::bind_cols(traction.kr)

spatialized.kr.sf = sf::st_as_sf(spatialized.kr, coords = c("Y", "X"))

plot(spatialized.kr.sf["traction.kr"])

spatialized.kr.sf = sf::st_set_crs(spatialized.kr.sf, 3812)
spatialized.kr.sf = sf::st_transform(spatialized.kr.sf, 4326)

# Building a grid of pixels of 10 m² that spatially matches our grid of centers. To do so, we must work with projected data (to define the cellsize)
interpolation.grid.pixels.sf = sf::st_sf(sf::st_make_grid(x = sf::st_transform(shapefile, crs = 3812),  cellsize = 10, what = "polygons"))

# adding the predictions to the grid squares object
spatialized.kr.sf = interpolation.grid.pixels.sf %>%
        dplyr::bind_cols(spatialized.kr.sf)

# transforming to 4326
spatialized.kr.sf = sf::st_transform(spatialized.kr.sf, 4326)

# intersecting with shapefile
spatialized.kr.sf =  sf::st_intersection(spatialized.kr.sf, shapefile)

# mapping 
mapview::mapview(spatialized.kr.sf["traction.kr"])

# to make the map responsive
responsiveness = "\'<meta name=\"viewport\" content=\"width=device-width, initial-scale=1.0\">\'"
  
# defining the color palette for traction
tractionPal = leaflet::colorNumeric(
  palette = "RdYlBu", #"RdBl",
  reverse = TRUE,
  domain = spatialized.kr.sf$traction.kr, #spatialized$response,
  na.color = "black"
)

traction.map = leaflet::leaflet(spatialized.kr.sf) %>%
    # basemaps
    leaflet::addProviderTiles(group = "Satellite",
      providers$Esri.WorldImagery,
      options = providerTileOptions(opacity = 1)
    ) %>%
    leaflet::addProviderTiles(group = "Stamen",
      providers$Stamen.Toner,
      options = providerTileOptions(opacity = 0.25)
    ) %>%
    # centering the map
    leaflet::fitBounds(sf::st_bbox(spatialized.kr.sf)[[1]],
      sf::st_bbox(spatialized.kr.sf)[[2]],
      sf::st_bbox(spatialized.kr.sf)[[3]],
      sf::st_bbox(spatialized.kr.sf)[[4]]
    ) %>%
    # adding layer control button
    leaflet::addLayersControl(
      baseGroups = c("Satellite","Stamen"),
     # overlayGroups = c("prediction"),
      overlayGroups = c("response"),
        options = layersControlOptions(collapsed = TRUE)
      ) %>%
    # fullscreen button
    leaflet.extras::addFullscreenControl() %>%
    # location button
    leaflet::addEasyButton(easyButton(
      icon = "fa-crosshairs", title = "Locate Me",
      onClick = JS("function(btn, map){ map.locate({setView: true}); }"))) %>%
    htmlwidgets::onRender(paste0("
      function(el, x) {
      $('head').append(",responsiveness,");
      }")
    ) %>%
    # admin boundaries
    # predictions
    leaflet::addPolygons(
      group = "response",
      color = "#444444", stroke = FALSE, weight = 1, smoothFactor = 0.8,
      opacity = 1.0, fillOpacity = 0.6,
      fillColor = ~tractionPal(traction.kr),
      label = ~ paste(
        "prediction:", format(round(spatialized.kr.sf$traction.kr, 2), nsmall = 2)),
      # popup = ~as.character(traction.kr),
      highlightOptions = highlightOptions(color = "white", weight = 2,
        bringToFront = TRUE)
    ) %>%
    #legend
    leaflet::addLegend(
      position = "bottomright", pal = tractionPal, values = ~traction.kr,
      title = "traction",
      group = "response",
      opacity = 1
    )


```


